### 2015-04-29 docs/prelim_analysis.Rmd
updated with table and plot for all sites

---
output: pdf_document
---
# Preliminary community synchrony analysis

@@

```{r load_libs, echo=FALSE, message=FALSE}
library(communitySynchrony)
library(plyr)
library(reshape2)
library(synchrony)
library(ggplot2)
```

```{r get_synchs, echo=TRUE, include=TRUE}
site <- "Kansas"
spp_list <- c("BOCU","BOHI","SCSC")
num_spp <- length(spp_list)
ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  ks_data <- rbind(ks_data, spp_data)
} #end species looping for raw data
ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row

@@

out <- get_comm_synchrony(ts_data = ks_data)
obs_synch <- round(as.numeric(out$pgr_synchrony[1]), 2)
exp_synch <- round(out$pgr_expected_synch_ind_flucts, 2)
exp_demo_effect <- exp_synch - obs_synch
print(out$pgr_synchrony)
print(out$pgr_expected_synch_ind_flucts)
```

@@

>>>>

---
output:
  pdf_document:
    fig_caption: yes
---

# Preliminary community synchrony analysis

@@

```{r load_libs, echo=FALSE, message=FALSE}
library(communitySynchrony)
library(plyr)
library(reshape2)
library(synchrony)
library(ggplot2)
library(xtable)
library(gridExtra)
library(RColorBrewer)
```

```{r get_all_data, echo=FALSE}
site <- "Kansas"
spp_list <- c("BOCU","BOHI","SCSC")
num_spp <- length(spp_list)
ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  ks_data <- rbind(ks_data, spp_data)
} #end species looping for raw data
ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row

@@

site <- "Idaho"
spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
num_spp <- length(spp_list)
id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  id_data <- rbind(id_data, spp_data)
} #end species looping for raw data
id_data <- id_data[2:nrow(id_data),] #remove first NA row
id_data <- subset(id_data, species!="ARTR") #take out the shrub

site <- "Montana"
spp_list <- sort(c("BOGR","HECO","PASM","POSE"))
num_spp <- length(spp_list)
mt_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  mt_data <- rbind(mt_data, spp_data)
} #end species looping for raw data
mt_data <- mt_data[2:nrow(mt_data),] #remove first NA row

site <- "NewMexico"
spp_list <- sort(c("BOER","SPFL"))
num_spp <- length(spp_list)
nm_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  nm_data <- rbind(nm_data, spp_data)
} #end species looping for raw data
nm_data <- nm_data[2:nrow(nm_data),] #remove first NA row

site <- "Arizona"
spp_list <- sort(c("BOER","BORO"))
num_spp <- length(spp_list)
az_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  az_data <- rbind(az_data, spp_data)
} #end species looping for raw data
az_data <- az_data[2:nrow(az_data),] #remove first NA row
```

```{r get_synchs, echo=FALSE, include=TRUE}
out <- get_comm_synchrony(ts_data = ks_data)
obs_synch <- round(as.numeric(out$pgr_synchrony[1]), 2)
exp_synch <- round(out$pgr_expected_synch_ind_flucts, 2)
exp_demo_effect <- exp_synch - obs_synch
print(out$pgr_synchrony)
print(out$pgr_expected_synch_ind_flucts)
```

@@

```{r get_all_synch, echo=FALSE, include=TRUE}
out_ks <- out
out_id <- get_comm_synchrony(ts_data = id_data)
out_mt <- get_comm_synchrony(ts_data = mt_data)
out_nm <- get_comm_synchrony(ts_data = nm_data)
out_az <- get_comm_synchrony(ts_data = az_data)
```

```{r synchrony_table, echo=FALSE, include=TRUE, results='asis'}
get_table_metrics <- function(output){
  tmp <- output
  exp_tmp <- round(tmp$pgr_expected_synch_ind_flucts, 2)
  obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
  demo_tmp <- exp_tmp - obs_tmp
  return(c(obs_tmp, exp_tmp, demo_tmp))
}

table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
                                  get_table_metrics(out_id),
                                  get_table_metrics(out_mt),
                                  get_table_metrics(out_nm),
                                  get_table_metrics(out_az)))
table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
table_data <- table_data[,c("Site", "V1", "V2", "V3")]
table_data$spp <- c(length(unique(ks_data$species)),
                    length(unique(id_data$species)),
                    length(unique(mt_data$species)),
                    length(unique(nm_data$species)),
                    length(unique(az_data$species)))
colnames(table_data) <- c("Site",
                          "$\\phi_{r}$", 
                          "$\\text{E}(\\phi_{r})$", 
                          "$\\text{E}(\\sigma_{d})$",
                          "Species richness")

synch_cap <- "Observed and expected synchrony among species' per capita growth rates and expected effect of demographic stochasticity on observed synchrony. Species richness values reflect the number of species analyzed from the community, not the actual richness."
print(xtable(table_data, caption = synch_cap),
      caption.placement="top",
      include.rownames = F, 
      sanitize.colnames.function = identity)
```

```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site."}
out_ks$percent_cover$site <- "Kansas"
out_id$percent_cover$site <- "Idaho"
out_mt$percent_cover$site <- "Montana"
out_nm$percent_cover$site <- "New Mexico"
out_az$percent_cover$site <- "Arizona"
cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
                    out_mt$percent_cover, out_nm$percent_cover, 
                    out_az$percent_cover)

out_ks$growth_rates$site <- "Kansas"
out_id$growth_rates$site <- "Idaho"
out_mt$growth_rates$site <- "Montana"
out_nm$growth_rates$site <- "New Mexico"
out_az$growth_rates$site <- "Arizona"
pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
                  out_mt$growth_rates, out_nm$growth_rates, 
                  out_az$growth_rates)
all_spp <- unique(pgr_data$species)
color_match <- data.frame(species = all_spp,
                          color = brewer.pal(11,"Paired"))
pgr_data <- merge(pgr_data, color_match, by = "species")
cover_data <- merge(cover_data, color_match, by="species")

cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
                geom_line()+
                geom_point()+
                facet_grid(site~., scales = "free_y")+
                guides(color=FALSE)+
                ylab("Mean Percent Cover")+
                xlab("Year")+
                scale_color_manual(values = brewer.pal(11,"Paired"))

pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
              geom_line()+
              geom_point()+
              facet_grid(site~., scales = "free_y")+
              guides(color=FALSE)+
              ylab("Per Capita Growth Rate")+
              xlab("Year")+
              scale_color_manual(values = brewer.pal(11,"Paired"))
full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
```
